Measuring directionality in double-beta decay and neutrino interactions with kiloton-scale scintillation detectors

### 2014-01-06 Direction_JINST_v7.tex
version with referee comments and one pass by Henry. --Lindley

The advantage of liquid scintillators for measurements in the
$\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
providing energy resolutions of $\sim$5\% \sigma_{e}$/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
two better than water Cherenkov detectors, the other developed
technology that can be economically scaled to these large
masses. However, this energy resolution is much poorer than other technologies being used for $0\nu\beta\beta$ searches: Ge detectors\cite{gerda2013}, Te bolometers\cite{Alessandria:2011rc}, tracking detectors\cite{SuperNEMO}, liquid Xe time projection chambers (TPCs)\cite{EXO2012} and high pressure gaseous Xe TPCs\cite{NEXTsipm}.

@@

Re-emission of absorbed photons in the scintillator
bulk volume and optical scattering, specifically Rayleigh scattering, have not yet been included by default. A test simulation shows that the effect of optical scattering is negligible. The reason is that the cutoff is very steep below 360~nm and almost no photons reach the sphere, so optical scattering makes no difference for short wavelengths. Above about 395~nm, the attenuation length is greater than 20~m so both scattering and absorption are not very likely and scattering is negligible. The intermediate region is rather small. A similar argument holds for re-emission. Scattering length measurements and discussions can be found in Ref.~\cite{Wurm:2010ad}.

>>>>

The advantage of liquid scintillators for measurements in the
$\sim$1~MeV range is their scalability from 1~ton to 1~kiloton while
providing energy resolutions of $\sigma(E)=\sim$5 \% $/\sqrt{E(MeV)}$\cite{kam2013,borexino}. This is roughly a factor of
two better than water Cherenkov detectors, the other developed
technology that can be economically scaled to these large
masses. However, this energy resolution is much poorer than other technologies being used for $0\nu\beta\beta$ searches: Ge detectors\cite{gerda2013}, Te bolometers\cite{Alessandria:2011rc}, tracking detectors\cite{SuperNEMO}, liquid Xe time projection chambers (TPCs)\cite{EXO2012} and high pressure gaseous Xe TPCs\cite{NEXTsipm}.

@@

Re-emission of absorbed photons in the scintillator
bulk volume and optical scattering, specifically Rayleigh scattering, have not yet been included by default. A test simulation shows that the effect of optical scattering is negligible. As shown in figure~\ref{scattplot}, scattering causes a small tail at longer times. The reason is that the cutoff is very steep below 360~nm and almost no photons reach the sphere, so optical scattering makes no difference for short wavelengths. Above about 395~nm, the attenuation length is greater than 20~m so both scattering and absorption are not very likely and scattering is negligible. The intermediate region is rather small. A similar argument holds for re-emission. Scattering length measurements and discussions can be found in Ref.~\cite{Wurm:2010ad}.

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{graphs2/Cerenkov_PEs_scattering_comparison.pdf}
        \caption[]{Photoelectron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV). The default simulation (black) is compared to the simulation with optical scattering turned on (red). Scattering causes a small tail to longer times as expected. \label{scattplot}}
        \end{center}
\end{figure}
